\chapter{Introduction} % (fold)
\label{cha:introduction}
Bipedal locomotion is a necessary capability to enable humanoid robots to navigate a wide range of environments required for most practical real-world applications. However, bipedal locomotion is a very challenging control problem due to the postural instability of the bipedal form. The act of bipedal walking is a cyclical repetition of a single gait cycle, a pattern of limb movement, which consists of a stance phase (foot is making contact with the ground) and a swing phase (foot is off the ground reaching for the next stance position). Bipedal gait can also be separated into the single and double support phases, when one or both feet are in contact with the ground, respectively. 

The most popular technique to achieve walking for bipedal robots thus far has been through control strategies based on the Zero-Moment Point (ZMP) criterion \cite{Vukobratovic:2004wy}. ZMP-stable trajectories are often computed off-line. Typically, these strategies \cite{Kajita:1997vr,Sugihara:2002kq} are energetically inefficient since they are actively trying to maintain balance by keeping the ZMP within the region of foot support. Furthermore, the resulting gait does not utilize the natural dynamics of the system and consequently does not look ``human-like''. 

A parallel research direction has considered taking advantage of the structure dynamics to achieve a stable gait cycle without active control effort. This pioneering research by McGeer \cite{McGeer:1990uk} introduced a unique class of legged robots known as passive dynamic walkers \cite{Collins:2005vp}. These robots are designed to walk on an inclined surface so that the passive mechanism is powered by gravity alone \cite{Spong:1999vk}. In addition to producing highly efficient walk, the gait patterns generated using this approach are much more human-like in comparison to ZMP-based control. However, passive dynamic walkers lack robustness to perturbations due to very narrow regions of attraction. 

In this thesis, a novel walking control strategy which has the potential to increase robustness and energetic efficiency is presented. To enable experimental validation with physical hardware, a 14 degree-of-freedom (DOF) lower body humanoid robot was developed. The ultimate goal is to validate the effectiveness and usefulness of the presented walking control strategy by testing it on the physical biped hardware. 

At the core of the proposed walking control strategy is the Foot Placement Estimator (FPE) algorithm which was previously developed for 2D bipedal robots \cite{Wight:2008vt}. The FPE algorithm determines where an unstable biped must step to restore balance. One of the key novelties of the FPE theory is the wide range of applicability of the results. It can be used to augment existing walking control strategies or be extended to form dynamically stable gait cycles \cite{Wight:2008ii}.

The solution to the FPE equation itself can be used as a recovery mechanism (i.e. in the face of a destabilizing disturbance) with existing ZMP-based strategies. Alternatively, it can be used to increase the narrow regions of attraction which plague minimally actuated passive dynamic walkers \cite{Goswami:1996gn,Asano:2000wi,Kuo:1999tn}.

\section{Contributions} % (fold)
\label{sec:contributions}

\begin{enumerate}
	\item \textbf{Electromechanical Design of 14 DOF Bipedal Robot} \\ 
	The development of a 14 DOF bipedal robot for the purposes of research in bipedal locomotion is presented. Drivetrain component selection is performed using gait estimation and accounting for actuator dynamics. The mechanical chassis is designed with anthropometric dimensioning and the weight distribution is manipulated to make the physical system easier to control.  \\

	\item \textbf{Rapid Prototyping Toolchain Development} \\ 
	A toolchain is developed to streamline the iterative design process by automatically generating dynamic simulations with 3D visualization directly from Computer-Aided-Design (CAD) software. It enables fast incremental changes to revise the mechanical design and immediately analyze the resulting behaviour in simulation. The toolchain is used to improve the initial design and compute forward dynamics for the 14 DOF bipedal robot. \\

	\item \textbf{Extension of FPE Algorithm to 3D} \\ 
	The 2D FPE algorithm is extended to form complete gait cycles for a 3D bipedal robot. The proposed algorithm selects a 2D plane in the chosen direction of motion and generates trajectories to produce a forward moving momentum along the plane. A whole body motion control framework coupled with a finite state machine is used to track the generated trajectories and form dynamically stable gait cycles. The efficacy of the proposed walking control strategy is demonstrated through side stepping and forward walking in dynamic simulations. \\

	\item \textbf{Experimental Validation on Physical Hardware} \\ 
	Dynamic simulation models are designed with a shared code base, allowing the same controller to target either a simulated plant or the constructed 14 DOF biped using hardware-in-the-loop (HIL). The validity of the motion control framework and actuator dynamics models are demonstrated on the physical hardware using HIL. \\

\end{enumerate}

% section contributions (end)

\section{Outline} % (fold)
\label{sec:outline}
The remainder of this thesis is organized as follows. Chapter \ref{cha:background} outlines existing walking control strategies in the robotics literature. While most control strategies use a dynamic measure of balance to form complete gait cycles, some newer approaches are similar to the strategy presented in this thesis. These control strategies aim to use foot placement as a means to form robust and energetically efficient gait for bipedal locomotion. 

The initial electromechanical design and development of a 14 DOF lower body humanoid robot (biped) is presented in Chapter \ref{cha:design}. Human gait trajectories are used as an approximate starting point for actuators and drivetrain component selection. While it is understood that using human gait as a reference is ambitious, this approach results in a conservative initial design by selecting components capable of human-speed gait, which is significantly faster than is achievable by most current humanoid designs. The forward dynamics and 3D visualization for the final prototype design is generated using the toolchain presented in Chapter \ref{cha:toolchain}. This is used to test control strategies for the biped in simulation prior to implementing it on the physical hardware.

One of the key challenges in developing physical hardware for bipedal locomotion research is the impact of small design changes on the overall stability of the system. A toolchain is developed in Chapter \ref{cha:toolchain} to help improve the initial design and ultimately produce a system with adequate drivetrain performance. The toolchain streamlines the process of designing a mechanical prototype in Computer-Aided-Design (CAD) software and immediately analyzing its behaviour in dynamic simulations.

The extension of the 2D FPE theory to the 3D case is presented in Chapter \ref{cha:simulations}. A motion control framework is presented in order to simultaneously use the FPE algorithm to achieve walking and maintain stability throughout the gait cycle. The 2D FPE theory and the proposed extension to 3D are verified through dynamic simulations. 

Experiments validating the actuator dynamics and motion control framework on the physical hardware are described in Chapter \ref{cha:experiments}. The actuator dynamics models are verified to match the physical DC motors through HIL simulations. The motion control framework used for extending the FPE theory to 3D is also verified on physical hardware using HIL.  

Lastly, conclusions and future work regarding the proposed walking control strategy and developed physical hardware are presented in Chapter \ref{cha:conclusion}. 
% section outline (end)

% chapter introduction (end)