\chapter{Related Work} % (fold) ===============================================
\label{cha:background}
Bipedal locomotion is one of the most active areas of research within humanoid robotics. This section presents a brief literature review on the design and gait synthesis for humanoid robotic applications. 



%==============================================================================
%   E L E C T R O M E C H A N I C A L   D E S I G N 
%==============================================================================



\section{Humanoid Electromechanical Design} % (fold) ==========================
\label{sec:related_electromechanical_design}
The electromechanical design of humanoid robots have undergone several generations of revisions. 


\subsection{Degrees of Freedom} % (fold) ======================================
\label{sub:related_degrees_of_freedom}
The selection of degrees-of-freedom (DOF) of a humanoid robot plays an essential role in the design and development phase. 
% subsection degrees_of_freedom (end) =========================================


\subsection{Mechanism Design} % (fold) ========================================
\label{sub:related_mechanism_design}
Linkage subsystems used to carry out motion is also an area where several approaches exist. 
% subsection mechanism_design (end) ===========================================


\subsection{Passivity-Based Designs} % (fold) =================================
\label{sub:related_passive_designs}
Perhaps the simplist approach to designing bipedal robots are those without any active power. 
% subsection passivity_based_designs (end) ====================================


\subsection{Actively-Powered Designs} % (fold) ================================
\label{sub:related_active_designs}
The most common approach to electromechanical design for bipedal robots is through the generation of active power for locomotion. 
% subsection actively_powered_designs (end) ===================================


% section humanoid_electromechanical_design (end) =============================



%======================================================================
%   W A L K I N G  C O N T R O L  &  G A I T  G E N E R A T I O N
%======================================================================



\section{Walking Control Strategies and Gait Generation} % (fold) =============
\label{sec:related_control_strategies}
There exists a wide range of control strategies to achieve bipedal locomotion in robotics literature to date. Some of the most popular strategies are briefly reviewed in this section.


\subsection{Zero-Moment Point} % (fold) =======================================
\label{sub:related_zero_moment_point}
The most popular techniques to achieve walking have been trajectory generation and control strategies based on the Zero-Moment Point (ZMP) criterion \cite{Vukobratovic:2004wy}. The ZMP defines a point on the ground where the forces acting on a biped do not produce a moment about the axes parallel to the ground plane. If the ZMP is kept within the region of foot support, the biped is stable. This stability criterion can be used to compute ZMP-stable trajectories offline that maximize the stability margin by maximizing the distance from the ZMP to the boundary of the support region. In the on-line phase, the stable trajectories are tracked online to execute walking \cite{HuangEtAlTRA2001}.  ZMP-based trajectories can also be generated on-line \cite{KajitaEtAlICRA2003,TakenakaEtAlIROS2009}. 

One of the biggest drawbacks of this approach is the resulting trajectory does not provide any strategy to respond to disturbances due to uneven terrain or unexpected forces. Typically, these strategies \cite{Kajita:1997vr,Kajita:2001fk,Sugihara:2002kq} are also energetically inefficient since they are constantly trying to maintain balance by keeping the ZMP point within the region of foot support. Furthermore, the resulting gait does not utilize the natural dynamics of the system and consequently does not look human-like. Despite the drawbacks, some of the most popular humanoid robots today utilize some variation of ZMP based feedback for bipedal locomotion. The most popular application is the famous Honda Asimo \cite{Hirai1998}. Newer bipeds available for research like Aldebaraan's Nao \cite{Gouaillier2006} robot also use ZMP-based walking control strategies. 
% subsection zero_moment_point (end) ==========================================


\subsection{Passive Dynamics} % (fold) ========================================
\label{sub:related_passive_dynamics}
An alternative approach, first proposed by McGeer \cite{McGeer:1990uk}, introduced a unique class of legged robots known as passive dynamic walkers \cite{Collins:2005vp}. This approach takes advantage of the natural dynamics of the biped structure, and is capable of maintaining a stable gait cycle without active control effort. Fully passive mechanisms walk on an inclined surface so that the mechanism is powered by gravity alone \cite{Spong:1999vk}. In addition to producing highly efficient walk, the gait patterns generated using this approach appear more human-like in comparison to ZMP-based control. However, passive dynamic walkers lack robustness to perturbations due to very narrow regions of attraction.

Thus far, majority of the research in passive dynamics has been restricted the very specific scenario of walking on an incline. This lacks the versatility that is required for humanoid robots if they are to ultimately navigate in most human environments (level ground). Further investigative research was helpful in characterizing the nature of a passive gait cycle \cite{Goswami:1996gn} in terms of stability and energies. This ultimately gave rise to new strategies which were aimed towards emulating the work done by gravity on an inclined slope \cite{Asano:2000wi}. These hybrid approaches known as virtual passive dynamics improved the versatility of the simple legged robots to walk on level ground with minimal actuation as a replacement for gravity \cite{Asano:2004tv}. 

While these minimally actuated bipeds have been shown to exhibit similar energetics as purely passive machines \cite{Asano:2004jp}, most research is typically restricted to 2D dynamics in the sagittal plane. The challenges of extending these simple models to 3D (i.e. incorporating lateral dynamics) lies in the unstable motion introduced by mismatch of roll velocity and contact conditions \cite{Kuo:1999tn}. This is particularly challenging since even small disturbances can be completely destabalizing for passive dynamic walkers. Although the minimal actuation strategy based on passive dynamics is more versatile, the main difference between the two approaches simply boils down the source of energy used for walking. Regardless of the source, the quantity of energy provided stay within a stable limit cycle depends only on the slope of the incline \cite{Goswami:1996gn}. Thus, stabilizing the dynamics in both sagittal and frontal planes is particularly challenging since both strategies share the same underlying weakness to perturbations. 
% subsection passive_dynamics (end) ===========================================


\subsection{Force/Impedance} % (fold) =========================================
\label{sub:related_force_impedance}
\Incomplete
% subsection force_impedance_control (end) ====================================


\subsection{Machine Learning} % (fold) ========================================
\label{sub:related_machine_learning}
\Incomplete
% subsection machine_learning_strategies (end) ================================


\subsection{Foot Placement} % (fold) ==========================================
\label{sub:related_foot_placement}
Recently, an alternative problem formulation focusing on restoring balance has been proposed. The Foot Placement Estimator (FPE), introduced by Dwight et al \cite{Wight:2008ii} formulates an approach to restore balance by controlling swing foot position during the gait cycle. By using the conservation of angular momentum, the FPE equation determines the location on the ground where the total energy of an unstable biped after swing foot impact is equal to the peak potential energy. If a step is taken before the FPE location, the post impact energy of the system causes the biped to fall over. Conversely, stepping beyond the FPE location on the ground causes the biped to fall back onto the hind leg.

The solution to the FPE equation itself can be used as a recovery mechanism (i.e. in the face of a destabilizing disturbance) with existing ZMP-based strategies. Alternatively, it can be used to increase the narrow regions of attraction which plague minimally actuated passive dynamic walkers \cite{Goswami:1996gn,Asano:2000wi,Kuo:1999tn}. The key concept here is that the FPE-based integration would require minimal joint actuation only to align a swing food appropriately to recover from a potential fall. As shown in \cite{Wight:2008ii,Wight:2008vt}, FPE can also be extended to form complete gait cycles to achieve dynamically stable walking. However, there are several key assumptions which are violated when attempting to implement this approach on a physical 3D robot. Namely, the theory presented in the derivations assume that the legs are massless and it only deals with the 2D dynamics in the sagittal plane.

The capture point concept, developed by Pratt et al. \cite{Pratt:2006vy}, is conceptually similar to the FPE. While the derivation of FPE is based on a simple compass biped model with fixed parameters, the capture point theory was derived using complex motion models which included using a flywheel body to control/offset any disturbances through the use of rotational inertia. Ultimately, the simplicity of the model allowed the FPE theory to be extended to complete gait cycles, while the work presented by Pratt et al simply solved the problem of lateral stabilization \cite{Wight:2008ii}.

Recently, a more comprehensive approach using the capture point for foot placement as a means to develop full walking control strategies has been proposed. De Boer \cite{DeBoer:2012wp} focused on the ground/foot interaction to develop a robust and energetically efficient walking control strategy for a force-controlled compliant lower-body biped. While this approach is philosophically similar to the idea behind FPE, there are several key differences. The approach presented in this thesis uses simple local controllers to form complete gait cycles and can be used on position-controlled joints without any complex actuation systems. The capturability framework demonstrated in \cite{Pratt01092012} used separate controllers for the swing and stance legs whereas this approach uses a single global differential kinematic resolution for whole body motion control.
% subsection foot_placement_strategies (end) ==================================


% section walking_control_strategies_&_gait_generation (end) ==================



%======================================================================
%   2 D  F O O T  P L A C E M E N T  E S T I M A T O R
%======================================================================




% chapter background (end) ====================================================